---
title: "Tribal FMAG - Data Contextualization"
author: "Colin T. Annand for RAD-SPAS"
date: "2024-05-13"
output:
  rmdformats::html_clean:
    highlight: haddock
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
    number_sections: TRUE
    thumbnails: FALSE
    code_folding: hide
    df_print: paged
    DT: datatable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
lapply(c("dplyr","tidyr","ggplot2", "plotly", "lubridate","tufte","kableExtra","scales","ggthemes","data.table", "rmdformats","readr"), FUN=library, character.only=TRUE, quietly=TRUE)
```

```{r fig.margin=FALSE, fig.width = 1, fig.height = 1,  warning=FALSE, message=FALSE, cache=TRUE, echo=FALSE}
knitr::include_graphics("fema-logo-blue.svg")
```

# Brief
- The FMAG Program Office is seeking assistance in determining a fire cost threshold (FCT) for Tribal Nations under the FMAG Program.
- Similarly to a recent policy decision in reducing the monetary threshold for major disaster declarations with Public Assistance from $250K to $100K for Tribal Nations.
- **PA Division Director Pesapane** recommended reaching out to us regarding this issue.

## Discussion Notes
- *5/15/2024 with Hart Gillespie and SPAS*
- Tribal Nations can only serve as a recipient after the **State** has met the FMAG threshold. 
- Are we (FEMA) applying a more stringent policy restriction than required by regulation?
- Can we remove the language prohibiting a **Tribal Nation** being a *recipient* without the State meeting the FMAG threshold
- It is **quite rare** (1-2 occurences) of a state **NOT** meeting the threshold. So the individual thresholds are the more stringent portion of policy. 
 - FMAG Funds are available, and then

## Goal:
- What a comparable threshold would be for **Tribal Nations** or should a threshold be waived, in respect to TN operating as a recipient for FMAG. 
- Tribal operations are usually around a 10th of the size of county/state operations, so even the 'lowest' *individual* FCT of $100,000 might be reduced to around $10,000. 
 

## Data
```{r data read in}
fida34787 <- read_csv("tables/FIDA_34787_PA_PDA .xlsx", skip=6)
threshold_fire_2024 <- readxl::read_xls("tables/CY24 FCT.xls", skip=1)
#PA_tribal <- read_csv("tables/PA_Tribal.csv")  #extraneous
```

```{r CTA functions, echo=FALSE, include=FALSE}
#for viz
.simpleCap <- function(x) {
    s <- strsplit(x, " ")[[1]]
    paste(toupper(substring(s, 1, 1)), substring(s, 2),
          sep = "", collapse = " ")
}

```


### Fire Cost Thresholds
 - Individual and Cumulative Levels
 - Either the greater of $100,000 or 3 times the *per capita threshold*.

```{r display FCT data, echo=FALSE}

threshold_fire_2024[,1:4] %>%
  mutate(`INDIVIDUAL THRESHOLD` = dollar(`INDIVIDUAL THRESHOLD`),
        `CUMULATIVE THRESHOLD` = dollar(`CUMULATIVE THRESHOLD`),
        `2020 Population` = comma(`2020 Population`)) %>%
  kable %>% 
    kable_styling("striped", full_width = F, font_size = 14) %>%
    scroll_box(width = "800px", height = "350px")
```
 
### FMAG Data
- **Dates: 1988 - 2022**
- There are only two Tribal Independent (Recipient and Applicant) FMAG declarations in this data set. 
- Additional data was garnered by text searching applicant names which matched the following strings: <br>"Tribe|Nation|Band|Native|Indian|Reserv|TDSA|OTSA|Pueblo|Choctaw|Navajo"<br> and manually removing incorrect matches. 
- These declarations are FMAGs where a Tribal Nation is operating as sub-recipient to a state. 

```{r}
### FMAG SPECIFIC data
fmag <- readxl::read_xlsx("tables/fmags.xlsx")
fmag <- data.table(fmag) 

fmag %>% #show data
  kable %>% 
    kable_styling("striped", full_width = F, font_size = 14) %>%
    scroll_box(width = "800px", height = "350px")
```

#### What number of applicants within this data is listed as Tribal (Applicant and Recipient)?

```{r}
table(ifelse(fmag$IS_TRIBAL==1, "Tribal", "Other"))
```
- This is a small number of tribes.
- Utilizing text searching of Applicant Names below to get a more complete list of tribal applicants (sub-recipients under states)

```{r}
#Two INCIDENTS - DR 2478 (2003-07), 2429 (2002-06)
fmag.tribe <- fmag[IS_TRIBAL==1]

fmag.tribe %>% #show data
  kable %>% 
    kable_styling("striped", full_width = F, font_size = 14) %>%
    scroll_box(width = "800px", height = "350px")
```

```{r}
#tribal text searching of applicant names
fmag.TAG <- fmag %>% 
  mutate(Tribe_TAG = 
           ifelse(grepl("Tribe|Nation|Band|Native|Indian|Reserv|TDSA|OTSA|Pueblo|Choctaw|Navajo",
                        APPLICANT_NAME, ignore.case = T)==TRUE, yes = "Tribe", no="Other")) 

fmag.TAG$Tribe_TAG[which(fmag.TAG$APPLICANT_NAME=="KY NATIONAL GUARD")] <- "Other"
```

What number of FMAGs include Tribes with this method?

```{r}
length(unique(fmag.TAG$DISASTER_NUMBER[fmag.TAG$Tribe_TAG=="Tribe"])) 
```

```{r}
fmag.TAG.tribe <- 
  fmag.TAG %>% 
  filter(Tribe_TAG == "Tribe")

fmag.TAG.tribe %>% 
  kable %>% #show data
    kable_styling("striped", full_width = F, font_size = 14) %>%
    scroll_box(width = "800px", height = "350px")
```
- This is a larger selection of tribes, and while these are not *independent applicants*, they provide a better sample than the 2 specific declarations we have. 


# Threshold Methods

- Below we attempt to find a **proportional threshold** for FMAGs involving Tribes.
- First a visualization of the Total FMAG Spending for Tribes (as selected above) and Others.

```{r}
# Aggregation of data to specific FMAG Events.
fmag.agg <- 
fmag %>% 
  mutate(Tribe_TAG = 
           ifelse(grepl("Tribe|Nation|Band|Native|Indian|Reserv|TDSA|OTSA|Pueblo|Choctaw|Navajo",
                        APPLICANT_NAME, ignore.case = T)==TRUE, yes = "Tribe", no="Other")) %>% 
    filter(APPLICANT_NAME != "KY NATIONAL GUARD") %>% 
  group_by(Tribe_TAG, DISASTER_NUMBER) %>% 
  summarise(STATE=STATE[1],
            APPLICANT_NAME = .simpleCap(tolower(APPLICANT_NAME[1])),
            Total_DR_Cost = sum(PROJECT_AMOUNT, na.rm=T),
            Num_Projects = n())

#join to FCT Threshold data
fmag.agg <- left_join(fmag.agg, threshold_fire_2024[,1:4], by="STATE")

#fix names
names(fmag.agg) <- gsub(x = names(fmag.agg), pattern=" ", replacement = "_")

#Create a "ratio" between total costs and thresholds.
fmag.agg <- fmag.agg %>% mutate(
  Ratio = Total_DR_Cost / CUMULATIVE_THRESHOLD)


cumulative_ratio <- mean(fmag.agg$Ratio,trim = .2)
#1.18 - so this is not useful for PDA...
# Need to get a PDA cost, rather than the threshold, becuase this is always exceeded. Can think about applying the same "over" ratio tho...
```

## Breakdown of FMAG Aggregated Data
```{r}
fmag.agg %>% 
  group_by(Tribe_TAG) %>% 
  summarise(
            Total_FMAGs = n(),
            Total_Spending = dollar(round(sum(Total_DR_Cost))),
            Highest_Cost = dollar(max(Total_DR_Cost)),
            Lowest_Cost = dollar(min(Total_DR_Cost))) %>% 
  kable %>% 
    kable_styling("striped", full_width = F, font_size = 14)
```


## Visualizing FMAG Incidents
 - Note, this chart is interactive if you opened the file in your *browser*. 
```{r echo=FALSE, warning=FALSE, fig.full.width=TRUE, fig.width=10, fig.height=6}
fmag.agg %>% 
  ggplot(aes(x=Tribe_TAG, y=Total_DR_Cost,
             color=Tribe_TAG,
             fill=Tribe_TAG))+
  geom_jitter(aes(size = Num_Projects,
                  text = paste(
    "FMAG:", DISASTER_NUMBER, "<br>",
    APPLICANT_NAME, "<br>",
    "Total:", dollar(round(Total_DR_Cost,0))
    )))+
  geom_boxplot(alpha=.25, outlier.shape = NA, color="black")+
  scale_y_log10(labels=dollar)+
  labs(title = "FMAG Disaster Level Costs",
       subtitle = "each dot a Disaster (aggregation of State or Tribe projects)",
       y= "$ Cost of Total Projects",
       x="",
       caption="Tribes identified by text search, not independent DRs") ->
  plot.fmag.DR.tribe

ggplotly(plot.fmag.DR.tribe, tooltip = "text")
```
- Each dot is an FM Incident (aggregation of all State or Tribe projects under one specific FMAG number). 
- States on the left in Red, Tribes on the right in blue. 
- The box gives the 'typical range' of FMAG incidents (obscured by the large number of points for States)
- In general, tribal FMAG activity looks like a 'scaled down' version of State/Terr activity. 
- Note: the lowest individual threshold of **$100,000** would screen out all but a few Tribal Nations if they were applying *independently* as opposed to as subrecipients under a state.


## Crafting a Threshold
- Using ratios of State Spending to the State Threshold (Cumulative)
- Calculating a 'Tribe Threshold' with that same ratio, giving a *proportional* number. 
```{r}
# Take the median (50%) DR costs for STATES. 
state_median <- median(fmag.agg$Total_DR_Cost[fmag.agg$Tribe_TAG!="Tribe"])
#'Median FMAG Cost for States is $614,573'

# Divide the median by the lowest Cumulative Threshold (could also use the median of State Thresholds...) $100,000
ratio <- 100000 / state_median

#1e06/state_median #the disaster threshold is set at %5, median of total costs across counties 
#0.049

tribal_median <- median(fmag.agg$Total_DR_Cost[fmag.agg$Tribe_TAG=="Tribe"])
#'Median FMAG Cost for Tribes is $139,811'

print(paste(
  "Using the ratio of threshold to median cost for FMAG decs, the lower bound for Tribal FMAG activity is",
dollar(tribal_median*ratio), "while the higher bound calculated using the Cumulative threshold is",
dollar(tribal_median*cumulative_ratio)))

```
- Median FMAG Cost for States is **$614,573**.
- Median FMAG Cost for Tribes is **$139,811**.
- Using a ratio from the *lowest* Individual Threshold, a *proportionate* Individual Threshold for Tribal nations is **$5,723**.
- Using a ratio derived from DR Cost / Cumulative Thresholds (per state) The *proportionate* Cumulative Threshold for Tribal Nations is **$13469.43**.

## Analytic Takeaway
There is a small slice of data specific to Tribal Nations within the FMAG universe. Any takeaway from this data contextualization should be appraised cautiously, however a general threshold range of:
- $5,700 to $13,500 
- Could be considered about proportional to the individual/cumulative State thresholds for FMAG funding. 
- While there are other ways to calculate the threshold, this range **illustrates** the low floor (as seen in specific Tribal FMAG activity). 
- Indeed, a threshold of 5,700 would actually have screened out at least one lower cost FMAG for a tribe (*Sycuan Indian Reservation* hover over the tribal points in the chart to see the values)



# Appendix
Further tables with costs, the distribution of projects, and futher data is listed below:

## Additional Plots

### Total Spending for Tribal FMAG
```{r}
fmag.agg %>% 
  group_by(Tribe_TAG) %>% 
  summarise(
            Total_FMAGs = n(),
            Total_Spending = dollar(round(sum(Total_DR_Cost))),
            Highest_Cost = dollar(max(Total_DR_Cost)),
            Lowest_Cost = dollar(min(Total_DR_Cost))) %>% 
  kable %>% 
    kable_styling("striped", full_width = F, font_size = 14)
```


### Distribution of Project Level Costs between Other and Tribes
```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=6}

fmag %>% 
  mutate(Tribe_TAG = 
           ifelse(grepl("Tribe|Nation|Band|Native|Indian|Reserv|TDSA|OTSA|Pueblo|Choctaw|Navajo",
                        APPLICANT_NAME, ignore.case = T)==TRUE, yes = "Tribe", no="Other")) %>%
  filter(APPLICANT_NAME != "KY NATIONAL GUARD") %>% 
  ggplot(aes(x=Tribe_TAG, y=PROJECT_AMOUNT,
             color=Tribe_TAG,
             fill=Tribe_TAG))+
  geom_jitter(aes(text = paste(
    APPLICANT_NAME, "<br>",
    "Total:", dollar(round(PROJECT_AMOUNT,0))
    )))+
  geom_boxplot(alpha=.75, outlier.shape = NA, color="black")+
  scale_y_log10(labels=dollar)+
  labs(title = "FEMA Project Level Costs",
       caption = "each dot a State or Tribe project",
       y= "$ Individual Project Costs",
       x="") -> 
  plot.fmag.proj.tribe

library(plotly)
ggplotly(plot.fmag.proj.tribe, tooltip = "text")

```

## Data Provided
-The following code takes the tables used in this analysis and converts them to a single .xlsx file for user reference:
```{r}
Tribal_FMAG_SPAS_data <- list(
  "FMAG_Applicants_1988-2022_raw" = fmag,
  "FMAG_Applicants_Tagged" = fmag.TAG,
  "FMAG_Tribal_Only" = fmag.TAG.tribe,
  "FMAG_Aggregated_Dec" = fmag.agg)

#writexl::write_xlsx(
#  x = Tribal_FMAG_SPAS_data, 
#  path =  "tables/Tribal_FMAG_SPAS_data.xlsx")
```


## openFEMA Data
- National Delivery Model 2017-Current
```{r}
decs <- readr::read_csv("DisasterDeclarationsSummaries (1).csv")

table(ifelse(decs$tribalRequest==1, "Tribe", "Other"), decs$declarationType)
```
- Open FEMA doesn't actually list Tribal Declarations under FMAG. This corresponds with the data above, where the **Independent** (Tribe as Applicant and Recipient) declarations were previous to 2003.
- These occured in Nevada, likely because Nevada was the sole state which **bore the cost of FMAG activity** instead of having the recipient do so, so most Tribes would prefer to apply under states. 


